### 2024-07-29 paper/main.tex
updated figures not sure what
started getting everything into latex

\documentclass[12pt,letterpaper]{article}
\usepackage{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{fontspec} %XeLatex
\usepackage{soul,xcolor}
\usepackage{graphicx,wrapfig}
\usepackage{float}
\usepackage{caption}
\graphicspath{ {./figures/} }
\usepackage[backend=biber]{biblatex}
\usepackage[paper=letterpaper,
            %includefoot, % Uncomment to put page number above margin
            marginparwidth=0.0in,     % Length of section titles
            marginparsep=.05in,       % Space between titles and text
            margin=1.0in,               % 1 inch margins
            includemp]{geometry}
\usepackage{fontspec} %XeLatex
\setmainfont{AvenirLTProBook.otf}[
    Path =fonts/,
    BoldFont = AvenirNextLTProBold.otf,
    %BoldFont = AvenirLTProHeavy.otf, % another option for bold font
    ItalicFont = AvenirLTProBookOblique.otf,
    BoldItalicFont = AvenirLTProHeavyOblique.otf
]
% Set caption styles
% \DeclareCaptionLabelFormat{table}{1-2}
\captionsetup[table]{labelfont=bf,labelsep=period}
% \captionsetup[figure]{labelfont=bf,labelsep=period}

\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays

\newcommand{\SethCaption}[2]{\noindent\textbf{\textcolor{black}{#1}} {#2}}
\newcommand{\SectionHeader}[1]{\noindent\textbf{\Large{\textcolor{black}{#1}}}\normalsize }
\newcommand{\SubSectionHeader}[1]{\noindent\Large{\textcolor{black}{#1}}\normalsize }

% \definecolor{myudblue}{RGB}{0, 83, 159}
\begin{document}
% \SectionHeader{Main Figures}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figures/exp1_experimental_design.png}

%     \SethCaption{Figure 1.}{Experimental Design}
% \end{figure}
% \begin{figure}[H]
%     \centering
%     \includegraphics[scale =1]{figures/exp1_movement_onset_panel.png}

%     \SethCaption{Figure 2.}{Data}
% \end{figure}
% \begin{figure}[H]
%     \centering
%     \includegraphics[scale = 1]{figures/exp1_score_metrics_panel.png}

%     \SethCaption{Figure 3.}{Data}
% \end{figure}

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{fontspec} %XeLatex
\usepackage{soul,xcolor}
\usepackage{graphicx,wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{arydshln} % For \hdashline
\usepackage{lineno}
\usepackage{setspace}

\definecolor{mydarkblue}{RGB}{0,51,102}
\graphicspath{ {./figures/} }

\usepackage[%
  autocite  = superscript,
  backend  = bibtex,
  sortcites  = false,
  citestyle = numeric,
  bibstyle=authoryear,
  sorting = none,
  maxbibnames=99,
  giveninits = true, %initials
  isbn=false,
  doi = false,
  url = false
]{biblatex}

\addbibresource{Aim1.bib}
% \bibliography{draft}

\usepackage[paper=letterpaper,
            %includefoot, % Uncomment to put page number above margin
            marginparwidth=0.0in,     % Length of section titles
            marginparsep=.05in,       % Space between titles and text
            margin=1.0in,               % 1 inch margins
            includemp]{geometry}

\renewbibmacro{in:}{}
\DeclareFieldFormat[article]{citetitle}{#1}
\DeclareFieldFormat[article]{title}{#1}  %
\DeclareFieldFormat{pages}{#1}% no prefix for the `pages` field in the bibliography
\DeclareNameAlias{sortname}{family-given} % put last name first
\DeclareNameAlias{default}{family-given}
\DeclareFieldFormat{labelnumberwidth}{\textcolor{mydarkblue}{\mkbibbold{#1\adddot}}} %remove brackets in bibliography
\DeclareFieldFormat{journaltitle}{\mkbibemph{#1}\isdot}
%\renewcommand{\labelnamepunct}{\addspace}  % remove period after year

% add numbers in front of citations and puts year behind authors
\defbibenvironment{bibliography}
  {\list
     {\printtext[labelnumberwidth]{%
    \printfield{prefixnumber}%
    \printfield{labelnumber}}}
     {\setlength{\labelwidth}{\labelnumberwidth}%
      \setlength{\leftmargin}{\labelwidth}%
      \setlength{\labelsep}{\biblabelsep}%
      \addtolength{\leftmargin}{\labelsep}%
      \setlength{\itemsep}{\bibitemsep}%
      \setlength{\parsep}{\bibparsep}}%
      \renewcommand*{\makelabel}[1]{\hss##1}}
  {\endlist}
  {\item}

\usepackage{xpatch}
\xpatchbibmacro{date+extrayear}{%
  \printtext[parens]%
}{%
  \setunit{\addperiod\space}%
  \printtext%
}{}{}
\AtEveryBibitem{%
  \clearfield{note}%
}

% adds comma after journal 
\renewcommand\bibpagespunct{\ifentrytype{article}{\addcolon}{\addcomma}\space}
\renewbibmacro*{volume+number+eid}{%
\setunit*{\addcomma\space}% NEW
  \printfield{volume}
  \printfield[parens]{number}{,}%
  \setunit{\addcomma\space}%
  \printfield{eid}}

% change and to \& (can also remove)
\DefineBibliographyExtras{english}{%
  \renewcommand*{\finalnamedelim}{\addcomma\addspace\&\addspace}%
  %\renewcommand*{\finalnamedelim}{\addcomma\addspace}%
}

%\usepackage{hyperref}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\setmainfont{AvenirLTProBook.otf}[
    Path =fonts/,
    BoldFont = AvenirNextLTProBold.otf,
    %BoldFont = AvenirLTProHeavy.otf, % another option for bold font
    ItalicFont = AvenirLTProBookOblique.otf,
    BoldItalicFont = AvenirLTProHeavyOblique.otf
]
% Set caption styles
% \DeclareCaptionLabelFormat{table}{1-2}
% \captionsetup[table]{labelfont=bf, labelsep=period}
\captionsetup[figure]{labelfont=bf, labelsep=colon, font=footnotesize, justification=justified}

\DeclareMathOperator*{\argmax}{srg\,max} % thin space, limits underneath in displays

\newcommand{\SuppCaption}[2]{\noindent\textbf{\textcolor{black}{#1}} {#2}}
\newcommand{\SectionHeader}[1]{\noindent\textbf{\Large{\textcolor{black}{#1}}}\normalsize }
\newcommand{\SubSectionHeader}[1]{\noindent\Large{\textcolor{black}{#1}}\normalsize }

\abovedisplayskip = 10.0pt plus 2.0pt minus 20.0pt
\belowdisplayskip = 10.0pt plus 2.0pt minus 20.0pt
\abovedisplayshortskip = 0.0pt plus 20.0pt
\belowdisplayshortskip = 9pt plus 3pt minus 20pt

% Condition variables
% \def\emlv{\emph{early mean low variance}}
% \def\emhv{\emph{early mean high variance}}
% \def\mmlv{\emph{middle mean low variance}}
% \def\mmhv{\emph{middle mean high variance}}
% \def\lmlv{\emph{late mean low variance}}
% \def\lmhv{\emph{late mean high variance}}

% \def\em{\emph{early mean}}
% \def\mm{\emph{middle mean}}
% \def\lm{\emph{late mean}}

% \definecolor{myudblue}{RGB}{0, 83, 159}
\begin{document}
%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%
\pagenumbering{gobble} % No page number for title page
\begin{center}

    \noindent\textbf{\Large\textcolor{mydarkblue}{Indecisions under time pressure arise from suboptimal switching behaviour
        }}
    \vspace{5mm}
    \\
    Seth R. Sullivan\textsuperscript{1}, \textsuperscript{2} Christopher Peters\textsuperscript{1}, \textsuperscript{2}, Rakshith Lokesh\textsuperscript{1}, Jan A. Calalo\textsuperscript{1},\\ Truc Ngo\textsuperscript{3}, John H. Buggeln\textsuperscript{5},Isaac Kurtzer\textsuperscript{5},Michael Carter\textsuperscript{5}, Joshua G.A. Cashaback\textsuperscript{1},\textsuperscript{2},\textsuperscript{4},\textsuperscript{5}
    \vspace{2mm}
    \\
\end{center}

\textsuperscript{1} Department of Biomedical Engineering, University of Delaware\\
\textsuperscript{2} Department of Mechanical Engineering, University of Delaware\\
\textsuperscript{3} Human Performance Lab, University of Calgary\\
\textsuperscript{4} Biomechanics and Movements Science Program, University of Delaware\\
\textsuperscript{5} Interdisciplinary Neuroscience Graduate Program, University of Delaware\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Abbreviated Title:}
\vspace{1mm}
\\
Indecisive behaviour arises from suboptimal switching
\\
\vspace{15mm}
\\
\textcolor{mydarkblue}{\large Funding:}
\vspace{1mm}
\\
XXXXXXXXXXXXXXX
\\
\vspace{10mm}
\\
\textcolor{mydarkblue}{\large Correspondence:}
\vspace{1mm}
\\
Seth Sullivan\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 122\\
Newark, DE 19711, U.S.A\\
Email: sethsullivan99@gmail.com
\vspace{2mm}
\\
Or
\vspace{2mm}
\\
Joshua G. A. Cashaback, PhD\\
Biomedical Engineering\\
University of Delaware\\
STAR Campus, Room 201J\\
Newark, DE 19711, U.S.A\\
Email: cashabackjga@gmail.com

\newpage
\doublespace %
\linenumbers %
\noindent \textbf{\large\textcolor{mydarkblue}{ABSTRACT}}
\vspace{2mm}
\\
Indecisive behaviour can be catastrophic, leading to car crashes or stock market losses. Despite fruitful efforts across several decades to understand decision-making, there has been little research on what leads to indecision. Here we examined how indecisions arise under high-pressure deadlines. In our first experiment, participants attempted to select a target by either reacting to a stimulus or guessing, when acting under a high pressure time constraint. We found that participants were suboptimal, displaying a below chance win percentage due to an excessive number of indecisions. Computational modeling suggested that participants were excessively indecisive because they failed to account for a time delay and temporal uncertainty when switching from reacting to guessing, which has not been previously reported in the literature. In a followup experiment, we show for the first time the existence of a time delay and temporal uncertainty when switching from reacting to guessing. Collectively, our results support the idea that humans are suboptimal and fail to account for a time delay and temporal uncertainty when switching from reacting to guessing, leading to indecisive behaviour.
\newpage

\noindent \textbf{\large\textcolor{mydarkblue}{NEW AND NOTEWORTHY}}
\vspace{2mm}
\\
The sensorimotor system has the constant challenge of dealing with the naturally occurring variability in our movements. Here we investigated the potential roles of muscular co-contraction and visuomotor feedback responses to regulate movement variability. When we visually amplified movements, we found that the sensorimotor system primarily uses muscular co-contraction to regulate movement variability. Interestingly, we found that muscular co-contraction was relative to participant-specific visuomotor feedback response, suggesting an interplay between impedance and feedback control.
\newpage
%\noindent\textbf{Author Summary}`
%\vspace{2mm}\\
%
%\newpage
\pagestyle{fancy}
%\cfoot[]{\textcolor{mydarkblue}{\thepage}}
\fancyhead[R]{\emph{\textcolor{mydarkblue}{Regulation of Movement Variability via Impedance and Feedback Control
        }}}
\fancyfoot[C]{\textcolor{mydarkblue}{\thepage}}
%\fancyfoot[C]{\emph{\textcolor{mydarkblue}{\thepage}}}
%\cfoot{\textcolor{myudblue}{\thepage}}

%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagenumbering{arabic} % page numbering starts here
\noindent\textbf{\large\textcolor{mydarkblue}{INTRODUCTION}}
\vspace{2mm}
\\
Indecisions often arise from failing to decide and act upon sensory information in time, such as a driver failing to brake or hit the gas pedal when a traffic light turns yellow. When acting under high pressure time constraints, the ability to accurately time a decision is critical to success. Past literature has had very little focus on indecisions. The vast majority of decision-making research either does not consider responses made after some time constraint or simply does not permit a non-response, such as in the classic two-alternative forced choice paradigm (Zachsenhouse Robust versus optimal,  Bogacz 2006, Cho et al. 2002 Mechanisms underlying, Jogan and Stocker 2014 A new two-alternative, Ratcliff et al. 2018 modeling 2-alternative forced-choice tasks, Tyler and Chen 2000, Ulrich and Miller 2004 Threshold estimation,   ). Thus, despite its real-world ubiquity and importance, we have very little understanding of how indecisions arise.

There have only been a handful of papers to examine indecisions, which involve either high (Lokesh et al. 2022) or low time pressure (Karsilar et al. 2014, Wu et al. 2016, Phillastides et al. 2011, Dambacher and Hubner 2015, Forstmann et al. 2008). We recently found a high proportion of indecisions during a competitive decision-making task between two humans that observed each other’s movements when selecting a target (Lokesh et al. 2022). In this competitive scenario, the ‘prey’ attempted to end up in the same target as the ‘predator’ by a time constraint, while the predator attempted to end up in the opposite target as the prey. This task had a high time pressure, such that participants were awarded no points if they were indecisive by failing to enter either target within the time constraint. The task poses a conundrum: it may be advantageous to wait for future sensory information and react to an opponent, but it could also be better to switch from reacting to guessing before the time constraint to avoid an indecision. Surprisingly, participants displayed a median indecision rate of approximately 25% with an upper range close to 40% indecisions. Similar to others (Wu et al. 2016, Phillastides et al. 2011, Dambacher and Hubner 2015, Forstmann et al. 2008), Karsilar and colleagues (2014) used a low time pressure task and found only 1.7% of trials were indecisions. Tasks with low time pressure are characterized by providing relatively strong sensory information well in advance of the time constraint deadline. Yet the mechanisms that give rise to indecisive behaviour, which is particularly relevant under common high time pressure scenarios, remains unclear.

Humans and animals attempt to maximize reward to time, select, and indicate a decision with a motor response (drugowitsch, balci, bogacz, hudson landy, trommershauser). To obtain more reward, it has been shown that it is important to consider the inherent time delays and temporal uncertainties of the nervous system (drugowitsch, bogacz, maybe balci too, hudson landy, acerbi wolpert internal representations, Wolpert/Faisal review, Kording/wolpert bayesian decision theory, wolpert/landy motor control is decision-making). Past work has shown that humans will often produce nearly optimal decision times during cognitive (Balci et al. 2011, Miletic and Van Maanen 2019) and motor tasks (Hudson et al. 2008, Faisal and Wolpert 2009 Near Optimal Combination). Other work has shown suboptimal action selection or timing, which has been suggested to occur from an imperfect representation of time delays or temporal uncertainties (Ota et al. 2015, drugowitsch computational precision). With time constraints, misrepresentations of inherent time delays or temporal uncertainties could lead to a missed deadline and consequently an indecision.

Building on our past work (Lokesh, 2022), we developed a high pressure task with a time constraint to examine how humans select decision times. We tested the idea that humans optimally account for time delays and temporal uncertainties to select a decision time that maximizes reward. Alternatively, humans may suboptimally represent time delays and temporal uncertainties, which can lead to an excessive number of indecisions. In Experiment 1, we found humans were suboptimal and observed excessive indecisions that led to a below chance win rate. Computational modelling work suggested that suboptimality arose by failing to account for the time delay and temporal uncertainty associated with switching from reacting to guessing. Experiment 2 showed for the first time, to our knowledge, the existence of an additional time delay and uncertainty when switching from reacting to guessing within a trial. Taken together, our work suggests that humans suboptimally represent the time delay and temporal uncertainty associated with switching from reacting to guessing, leading to indecisive behaviour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\noindent\textbf{\large\textcolor{mydarkblue}{RESULTS}}
\vspace{2mm}
\\
\textbf{\textcolor{mydarkblue}{{Experiment 1}}}
\vspace{2mm}

\noindent\textbf{\textcolor{mydarkblue}{{Experimental Design}}}
\vspace{2mm}

The goal of Experiment 1 was to test how stimulus timing influenced indecisive behaviour. Briefly, participants began each trial by moving their cursor into a start position (Fig. 1A). The stimulus, represented as a cursor on the screen, would quickly move to one of the two target circles. Participants were instructed to reach the same target as the stimulus within a time constraint of 1500 ms. The time remaining in each trial was represented visually with a timing bar that decreased in width according to the elapsed time. Thus, participants were fully aware of how much time they had left relative to the time constraint. A trial was considered a win and the participant received one point if they successfully reached the same target as the stimulus within the time constraint (Fig. 1B). A trial was considered incorrect and the participant received zero points if they reached the opposite target as the stimulus within the time constraint. A trial was considered an indecision and the participant received zero points if they failed to reach a target within the time constraint.

For each trial within a condition, the stimulus movement onset was drawn from the same normal distribution. Using a 3 x 2 within experimental design (Fig. 1C), in separate blocks we manipulated the stimulus movement onset mean (early mean = 1000 ms, middle mean = 1100 ms, late mean = 1200 ms) and standard deviation (low variance = 50 ms, high variance = 150 ms). For the purposes of the main manuscript we focus on the results of the low variance conditions, but report the findings for the high variance conditions in Supplementary A.

\noindent\textbf{\textcolor{mydarkblue}{{Participant Timing Behaviour}}}
\vspace{2mm}

Participant movement onsets for the low variance condition are shown in Fig. 2A, We found a significant main effect of stimulus movement onset mean (F[1.551,29.475] = 4.36, p = 0.030) and variance (F[1.000,19.000], p = 0.017). There was no significant interaction between stimulus movement onset mean and variance (F[1.657,31.474], p = 0.565). When collapsed across low and high variance, participant movement onsets were significantly greater in the middle mean conditions compared to the early mean conditions (p = 0.014, $\hat{\theta}$=72.5 \%), suggesting that participants waited longer to react to the stimulus movement and guessed later in time. Again when collapsed across low and high variance, participant movement onset significantly decreased from the middle mean conditions to the late mean conditions (p = 0.018, $\hat{\theta}$=62.5\%). Here, an earlier participant movement onset in the late mean condition suggests that participants attempted to wait and react to the stimulus, but ended up guessing.

Participant movement onset standard deviation for the low variance conditions are shown in Fig. 2B. There was a main effect of mean (F[1.383, 26.284], p = 0.018) and variance (F[1.000,19.000], p < 0.001) of the stimulus movement onset, and no significant interaction (F[1.978, 37.575], p = 0.097). When collapsed across variance, waiting to react and then guessing in the late mean conditions led to a higher standard deviation of participant movement onset relative to the middle mean (p = 0.039, $\hat{\theta}$ = 70.0\%) and early mean (p < 0.001, $\hat{\theta}$ = 77.5\%) conditions.

\noindent\textbf{\textcolor{mydarkblue}{{Participants are suboptimal and excessively indecisive.}}}

We calculated the indecisions (Fig. 2C), wins (Fig. 2D), and incorrects (Fig. 2E) for each of the experimental conditions. Participants displayed a high proportion of indecisions. The median percentage of indecisions was 15.0\% [range: 0.0 - 93.8\%] across all conditions, with the late mean low variance condition having a median percentage of indecisions of 19.4\% [range: 1.2\%, 93.8\%]. We found a significant interaction between stimulus movement onset and variance for indecisions (F[1.571, 28.781] = 5.58, p = 0.013). In low variance conditions, participants made significantly more indecisions in the middle mean condition than the early mean condition (p<0.001, $\hat{\theta}$ = 85.0\%; Fig. 2C). Additionally, participants made significantly more indecisions in the late mean condition compared to the early mean condition (p<0.004, $\hat{\theta}$ = 80.0\%).

The win percentage across all conditions was 56.25\% (range: 6.2\%, 93.8\%; Fig. 2D]. We found a significant interaction between stimulus movement onset mean and variance for wins (F[1.542, 29.296] = 23.73, p<0.001). We found that the late mean condition had significantly less wins than the early mean condition (p < 0.001, $\hat{\theta}$ = 95.0\%). Interestingly, in the late mean condition we found that the average win percentage was significantly below the 50\% chance level (p<0.001; $\hat{\theta}$=95.0\%), which was true for 19 out of 20 participants. Since guessing on every trial would lead to a win percentage of 50\%, the only way participants would be below chance is if they are excessively indecisive.
The incorrect percentage across all conditions was 26.3\% [range: 0.0\%, 57.5\%; Fig. 2E]. We found significant interactions between stimulus movement onset mean and variance for incorrects (F[1.658,31.508]=3.72, p=0.033). Participants displayed a greater percentage of incorrect decisions in the late mean condition than the early mean condition (p < 0.001, $\hat{\theta}$ = 92.5\%).

\noindent\textbf{\textcolor{mydarkblue}{{Decision making model.}}}

In our task, participants must make a decision of whether to react to the stimulus or guess. For Experiment 1, we tested three different models:  i) No Switch Time Model, ii) Full Switch Time Model and iii) Partial Switch Time Model (Fig. 3, left column). The decision policy of these models considers the expected value ($\mathbb{E[R|\tau]}$) to determine the time ($\tau$) to transition from reacting to guessing. Expected value is defined as

\begin{align}
    \mathbb{E}[R|\tau] = & P(Win|\tau) \cdot R_{Win} \nonumber \\ &+ P(Incorrect|\tau) \cdot R_{Incorrect} \nonumber \\ &+ P(Indecision|\tau) \cdot R_{Indecision}
\end{align}

where ($P(Win|\tau)$) is the probability of a win, ($P(Incorrect|\tau)$) probability of an incorrect, and ($P(Indecision|\tau)$) is the probability of an indecision. $R_{Win} = 1$, $R_{Incorrect} = 0$, and ($R_{Indecision} = 0$ correspond to the reward structure of the task (Fig. 1B). The decision policy of each model maximized expected reward to determine the optimal time to transition from reacting to guessing, $\tau^*$, according to

\begin{equation}
    \tau^* = \underset{\tau}{argmax}[\mathbb{E}(R|\tau)]
\end{equation}

Each model has varying knowledge of the different parameters (Fig. 3, left column). A model can have full knowledge or partial knowledge of a particular parameter. With full knowledge, the decision policy fully utilizes the parameter when selecting the time to transition from reacting to guessing. With partial knowledge, the decision policy utilizes its partial and imperfect representation of the parameter. Here the idea is that a human may not have a perfect representation of some parameter when determining a transition time, even though that particular parameter will still influence behaviour. All model parameter values are shown in Supplementary B.

\noindent\emph{\textcolor{mydarkblue}{{No Switch Time Model}}}

We first considered a model that incorporated various time delays and temporal uncertainties from sources previously identified in the literature: response time, neuromechanical delay, movement time, stimulus movement onset, and timing uncertainty. Note, unlike the other models we will address below, this model did not consider a `switch time’ delay and uncertainty because it was not considered in past literature. Hence, we termed it the No Switch Time Model.

In the late mean low variance condition, the No Switch Time Model underestimated participant movement onset (Fig. 2A), underestimated indecisions (Fig. 2C), and overestimated wins (Fig. 2D). During this condition, participants displayed 19\% indecisions on average and a win percentage significantly below chance. One reason that the No Switch Time Model was unable to capture behaviour is because it did not consider the potential delays and uncertainties that might exist when switching from reacting to guessing.

\noindent\emph{\textcolor{mydarkblue}{{Full Switch Time Model}}}

Next we considered a model that additionally incorporated the potential existence of a switch time delay and uncertainty when transitioning from reacting to guessing. For this Full Switch Time Model, we assumed that the model had full knowledge of the time delay and uncertainty when switching from reacting to guessing.

Yet, despite including switch time, the Full Switch Time Model also performed poorly when predicting participant movement onset (Fig. 2A), indecisions (Fig. 2C), and wins (Fig. 2D) in the late mean condition. An explanation for why this model did not do well to explain indecisions is that humans may not have full knowledge of this potential switch time delay and uncertainty.

\noindent\emph{\textcolor{mydarkblue}{{Full Switch Time Model}}}
Finally, we considered a model that had only partial knowledge of a potential switch time delay and uncertainty when transitioning from reaching to guessing. That is, this model specifically tests whether humans have an imperfect representation of a switch time delay and uncertainty. The model also had partial knowledge of timing uncertainty, which the fitting procedure found to further improve model fits. The Partial Switch Time Model was able to replicate all aspects of behaviour (Fig. 2). Crucially, it was also able to capture suboptimal behaviour in the late mean condition, where we found that an excessive percentage of indecisions (Fig. 2C, Fig. 3) led to a lower than chance win percentage (Fig. 2D).

\textbf{\textcolor{mydarkblue}{{Experiment 1}}}
\vspace{2mm}

Our behavioural findings in Experiment 1 demonstrated that participants were suboptimal decision makers. Through our modelling efforts, we were able to capture this suboptimal decision making by including a switch time delay and uncertainty when transitioning from reacting to guessing. The switch time delay and uncertainty were only partially represented by the Partial Switch Time Model when determining the optimal time to switch from reacting and guessing. However, we are not aware of any work that considers the temporal time delays and uncertainty associated with switching from reacting to guessing within a trial. Thus, the goal of Experiment 2 was to determine if there is indeed a switch delay and uncertainty that occurs when humans transition from reacting to guessing.

\noindent\textbf{\textcolor{mydarkblue}{{Experimental Design.}}}
For all conditions, participants controlled a visible cursor that was aligned with their hand position. They started each trial by moving their cursor into a start position. Trial onset began with the appearance of both the stimulus (yellow cursor) and two targets. Participants could experience two trial types: react trials or guess trials. In the react trials, participants saw the stimulus move and were instructed to as quickly as possible follow the stimulus to one of the targets (Fig. 4A). In the guess trials, participants saw the stimulus disappear from the start circle. They were instructed to guess one of the two targets as quickly as possible (Fig. 4B). Following trial onset, the movement or disappearance of the stimulus was drawn from a normal distribution with a mean of 800ms and a standard deviation of 50ms. There were three experimental conditions (Fig. 4C): the react or guess condition, the only react condition, and the only guess condition. In the react or guess condition, react trials and guess trials were randomly interleaved (50 react trials and 50 guess trials). Participants were informed that the stimulus would either move to one of the targets or disappear. In the only react condition, participants were informed that the stimulus would always move to one of the two targets (50 react trials and 0 guess trials). They were also told that the stimulus would not disappear. In the only guess condition, participants were informed that the stimulus would always disappear (0 react trials and 50 guess trials). They were also informed that it would not move to one of the two targets.

During the react or guess condition, we reasoned that participants would prefer to react because they would be guaranteed to select the correct target. As a result, in the react or guess condition, if the stimulus disappeared the participant would switch from reacting to guessing when selecting a target. Conversely, during the only guess condition, if the stimulus disappeared participants would not have to switch from reacting to guessing. Thus, if there is a delay when switching from reacting to guessing, we would expect a greater response time for the guess trials in the react or guess condition compared to the guess trials in the guess only condition.

\textbf{Response Time}. Average participant response times are shown for react and guess trials for each condition are shown in Fig. 5A. As expected, we found significantly greater response times for guess trials in the react or guess condition when compared to the only guess condition (p < 0.001, $\hat{\theta}$ = XX.X), which was displayed by all participants. These comparatively greater response times for guess trials in the react or guess condition supports the idea that there is a switch time delay when transitioning from reacting to guessing.

Likewise, if there was a switch time delay we would also expect a comparatively greater response time difference between guess and react trials in the react or guess condition, compared to the response time difference between only guess trials and only react trials [i.e., guess - react (react or guess condition) > guess - react (guess only and react only conditions)]. Indeed, we found a greater response time difference between guess and react trials in the react or guess condition, compared to the response time difference between the guess only and react only conditions (p < 0.001, $\hat{\theta}$ = XX.X, Fig. 5A). Moreover this result shows that the response time differences between guess and react trials are not due to any dual tasking (Selst and Jolicoeur 1997) or task switching between trials (Monsell 2003, Kiesel et al. 2010, Rubinstein 2001), which would not show this relative difference  [i.e., guess - react (react or guess condition) = guess - react (guess only and react only conditions)].

\textbf{Response Time Uncertainty}. We also examined participant response time uncertainty, calculated as the standard deviation (Fig. 5B). We found that response time uncertainty on guess trials was significantly greater in the react or guess condition compared to the only guess condition (p < 0.001). This result suggests that there is additional uncertainty when participants switch from reacting to guessing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\textbf{\large\textcolor{mydarkblue}{Discussion}}

We found that participants were suboptimal decision-makers and excessively indecisive, leading to a below chance win rate. Computational modelling suggested that excessive indecisions were a result of failing to account for a delay and uncertainty associated with switching from reacting to guessing. We then showed empirical evidence of an additional delay and uncertainty when switching from reacting to guessing. Taken together, we found that participants were suboptimal decision-makers and excessively indecisive because they did not account for the time delay and temporal uncertainty when switching from reacting to guessing.

In Experiment 1, participants were required to reach the same target as a cursor before a time constraint. Within a trial, they could either react to a moving stimulus or guess which of the two targets would be correct. We saw that 95\% of participants had a win rate less than chance (50\%) in the late mean condition, which corresponded with an average of 19.4\% indecisions. This proportion of indecisions aligns with our recent prior work that examined competitive human-human decision making with a high time pressure (Lokesh et al. 2022). In this competitive task, one participant attempted to reach the same target as their opponent, while the other tried to reach the opposite target within a time constraint. It was suggested that the high proportion of indecisions in this competitive human-human task were the result of participants waiting too long to acquire sensory information of their opponent, despite the impending time deadline. Likewise, we found a high proportion of indecisions across experimental conditions. Our results would also suggest that participants waited to acquire sensory information of when the stimuli would move. Moreover, building upon the previous work (Lokesh et al. 2022), our work also suggests that a key contributor leading to excessive indecisions is failing to account for the time delay and temporal uncertainty when switching from reacting to guessing.

Past work has suggested that humans can nearly optimally account for time delays and temporal uncertainty when performing decision making (Balci Risk assessment in man and mouse, Jazayeri and Shadlen 2010. Balci Optimal temporal risk assessment) and movement tasks (Hudson et al. 2008, Battaglia and Schrater 2007, Dean 2007) when attempting to maximize reward. Here we considered two optimal models, the No Switch Time Model and Full Switch Time Model, which both had full knowledge of the inputted time delays and temporal uncertainties. Interestingly, both the No Switch Time Model and Full Switch Time Model showed that even when fully accounting for all sensorimotor delays and uncertainties, indecisions were a part of an optimal strategy in all but one of the six conditions. In other words, given the inherent delays and uncertainty of our nervous system (Wolpert Faisal), an optimal strategy of earning maximal reward may involve indecisive behaviour on some proportion of trials. We are unaware of any work in the literature suggesting that some level of indecisions may be optimal. Even though making some indecisions can be optimal, our results in Experiment 1 were in support of the idea that humans are suboptimal. Specifically, in the late mean condition, we found that humans were suboptimal since they had a win percentage lower than chance, which arose from an excessive number of indecisions. The Partial Switch Time Model was suboptimal, since it had a partial representation of time delays and temporal uncertainties associated with switching from reacting to guessing. We found that this model best explained behaviour, including a below chance win percentage and an excessive number of indecisions. The Partial Switch Time Model supports the notion that humans suboptimally select decision times when under high time pressures. An interesting future direction would be to test whether different reward structures, such as placing a higher reward on wins or punishing indecisions (Kahneman and Tversky 2013, Roth et al. 2024, Galea et al. 2015), would provide a means to reduce an excessive number of indecisions.
The combined empirical evidence of Experiment 1 and computational modelling suggested the existence of time delay and uncertainty when switching from reacting to guessing. However, we were unaware of any work in the literature to support this idea. In Experiment 2 we tested the notion of a more delayed and uncertain response time when switching from reacting to guessing, compared to guessing by itself. Indeed, we found that when participants had to switch from reacting to guessing, their response times were significantly slower and more uncertain than when they only had to guess. One possibility for increased time delays and temporal uncertainty could be related to switching between different processing ‘modes’. In our task, participants may have switched from a `react mode’ that corresponded to preparing to follow the stimulus, and then switch to a `guess mode’ to randomly select one of the targets.

Dutilh and colleagues (2011) explored the idea of switching between a stimulus controlled (i.e. react) mode and a guess mode between trials. In their task, participants were required to discriminate between a word stimuli from a non-word stimuli by selecting one of two buttons during a two-alternative forced choice task. Between trials, the authors manipulated whether participants received more reward for fast decisions to promote guessing, or more reward for accurate decisions that promoted reacting. Participants displayed longer response times when transitioning from more accurate decisions to fast decisions, compared to when transitioning from fast decisions to accurate decisions given the same current reward weighting. The authors interpreted these results to represent a resistance, termed hysteresis, when switching between react and guess modes. That is, participants are more likely to stay in their current mode than switch modes. They also highlighted that classical decision-making models, such as drift-diffusion models (Bogacz, Ratcliff) or more recently the urgency-gating model (Cisek, Thura,derosiere thura cisek duque), do not consider different modes of reacting or guessing. Extending upon the findings of Dutilh and colleagues (2011) that examined between trial mode switching, our work suggests that humans switch between react and guess modes within a trial. Importantly, we find that not accounting for the time delays and temporal uncertainties when switching from reacting to guessing gives rise to excessive indecisions. To our knowledge, it is unknown how different modes would be represented in the nervous system. One possibility is that the different modes represent different attractors from a neural dynamical systems perspective (Erlhagen and Schoner 2002, wang 2008 Decision making in recurrent, Churchland et al. 2012, Shenoy Sahani Churchland 2013), which would be an interesting avenue of investigation.

We found in Experiment 2 that participant response times were more delayed by approximately 75ms during guess trials in the react or guess condition, compared to guess only trials. Moreover, the response time difference between guess and react trials in the react or guess condition are significantly greater than the response time difference between the guess only and react only conditions. Collectively, these results suggest that participants have an initial preference to react, before having to switch to a guess in the react or guess condition. In this experiment the behaviour may be explained by a strong preference to react since it yields a 100\% success probability, as opposed to guessing that on average yields a 50\% success probability. It would be interesting for future studies to examine if they can manipulate the magnitude or probability of reward to switch a preference between reacting or guessing, and how this impacts indecisive behaviour.
